---
title: "Basics of the 'trafficSafety' Package"
author: "Bojan Zunar"
date: "`r format(Sys.time(), '%d %m %Y')`"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{basics}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
wd <- dirname(system.file("extdata", "accident_2013.csv.bz2", package = "trafficSafety"))
```

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)

knitr::opts_knit$set(root.dir = wd)
```

Package 'trafficSafety' was primarily created for the Coursera's 'Building R Packages' course. It contains a number of functions that load, summarise, and visualise data from the US National Highway Traffic Safety Administration's 'Fatality Analysis Reporting System', which is a nationwide census providing the American public yearly data regarding fatal injuries suffered in motor vehicle traffic crashes.

# Loading the Dataset

The package comes with three files that list fatalities throughout the years 2013-2015. They can be loaded with function `fars_read` which outputs a tibble. For the convenience, filenames can also be generated through a helper function `make_filename`.

```{r loadData}
library(trafficSafety)

# external data bundled with the package
y13 <- system.file("extdata", "accident_2013.csv.bz2", package = "trafficSafety")
y14 <- system.file("extdata", "accident_2014.csv.bz2", package = "trafficSafety")
y15 <- system.file("extdata", "accident_2015.csv.bz2", package = "trafficSafety")

# load the data
fars13 <- fars_read(y13)
head(fars13, 3)

# helper function
make_filename(2013)
system.file("extdata", make_filename(2013), package = "trafficSafety")
```

Several years worth of data can also be imported using the function `fars_read_years` which will extract only months and year, saving data from each file as a separate element of the list.

```{r manyYears}
fars_read_years(c(2013, 2014))
```

# Summarise Number of Accidents

One can summarise the number of accidents by using `fars_summarize_years` function.

```{r summarise}
fars_summarize_years(c(2013, 2014, 2015))
```

# Visualise Loaction of the Accidents

Finally, one can visualise where the accidents occured for a specific state in the specified year using the `fars_map_state` function.

```{r visualise, fig.align = "center"}
fars_map_state(10, 2014)
```

```{r sessionInfo}
sessionInfo()
```

